---
title: "Port folio 4"
author: "Martine Lind Jensen"
date: "20/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(pacman)
pacman::p_load(MuMIn, tidyverse, reshape2)
```

```{r}
#loading data 
df <- read.csv("WHO_suicide_statistics.csv")
```

Checking for collinearity ##Dont know man


```{r}
#Corfunction
round(cor(df),2)

'#plot as heatmap
qplot(x=Var1, y=Var2, data=melt(cor(df)), 
      fill=value, 
      geom="tile") +
   scale_fill_gradient2(limits=c(-1, 1))


##### Highlight only the "worrying"" correlation values
#plot only values above .7
plot_df <- melt(cor(child))
plot_df$value[plot_df$value < 0.7 & plot_df$value > - 0.7] = 0
qplot(x=Var1, y=Var2, data=plot_df, fill=value, geom="tile") +
   scale_fill_gradient2(limits=c(-1, 1))'

```

Making coloumn that shows the suicide rate pr. population 

Making it promille??
```{r}
df$suicide_population <- (df$suicides_no/df$population)*1000
```


Running anova 
```{r}
#running anova and storing the output
anova_df <- aov(df$suicide_population ~ df$age, data = df)

#looking at the output
summary(anova_df)

#Running anova on the log data 

anova_log <- aov(df_log$log_suicide ~ df_log$age, data = df_log)

summary(anova_log)
```
There is a difference between the means, so we reject the null-hypothesis. 




Checking which agegroup is the one that stands out
```{r}

#On normal data
m1 <-  lm(df$suicide_population ~ df$age, data = df)

summary(m1)

#On log data 

mlog <- lm(df_log$log_suicide ~ df_log$age, data = df_log)

summary(mlog)
```
So this shows that 05-14 years(intercept) is not significantly different from 0 or the mean, pretty sure its 0??? Reject the null-hypothesis?

With the change of the suicide pr population promille the 05-14 is actually significant.

log transformation
```{r}
#Removing zeros, so we can do the log transformation. Need to argue that we can do that 
df_clean <- filter(df, df$suicides_no != "0")

#Making the log transformation
df_log <- df_clean %>% mutate(log_suicide = log(df_clean$suicide_population))

```


Trying to 
```{r}
m2 <- lm(df_log$log_suicide ~ df_log$age, data = df_log)

summary(m2)
```

Posthoc test
1) Benferroni correction:  conservative, controls the family wise error 

 "The Bonferroni adjustment simply divides the Type I error rate (.05) by the number of tests (in this case, three). Hence, this method is often considered overly conservative. The Bonferroni adjustment can be made using p.adj = “bonferroni” in the pairwise.t.test() function."

```{r}
#difference is present -> small p-values
pairwise.t.test(df_log$log_suicide, df_log$age, p.adjust.method = "bonferroni")

```
2) Tukey’s HSD correction: less conservative, better compromise between Type 1 and Type 2 error
  
  "The Tukey Honest Significant Difference (HSD) method controls for the Type I error rate across multiple comparisons and is generally considered an acceptable technique. This method can be executed using the multcomp::glht(Model1, linfct = multcomp::mcp(predictor = “Tukey”))."
  
```{r}
'#we need to make sure predictor is a factor
happy_allhigh$country <- as.factor(happy_allhigh$country)
happy_low$country <- as.factor(happy_low$country)

#we also need multcomp library
pacman::p_load(multcomp)

#no difference -> all p-values are very large
summary(glht(aov(happiness ~ country, data = happy_allhigh), linfct = mcp(country = "Tukey")))

##difference is present -> small p-values
summary(glht(aov(happiness ~ country, data = happy_low), linfct = mcp(country = "Tukey")))'
```


Check out this dataframe showing the suicides in these ages.
```{r}
#df_0514 <- filter(df, df$age == "05-14 years")
```


15-24 have 133.049

#### Checking assumptions 

1. Homogeneity of variance 

The assumption of homogeneity of variance is that the variance within each of the populations is equal. ANOVA works well even when this assumption is violated except in the case where there are unequal numbers of subjects in the various groups.

Can be checked using barlett.test(outcome ~ predictor, data = mydata):
  If p value < 0.5 - Significant result, therefore variances cannot be assumed to be equal (violated assumption)

  If p value > 0.5 - Non-significant result, therefore variances can be assumed to be equal (assumption is not violated)

```{r}

#Using the bartlett.test because that is what anita did 
bartlett.test(df_log$log_suicide ~ df_log$age, data = df)

#Using levenesTest 
car::leveneTest(df_log$log_suicide ~ df_log$age, data = df)


```

In this case, the result is non-significant (p-value > 0.05), meaning variances can not be assumed to be equal.



2. Checking residuals

plot(model): check everything at once
plot(model, 1): check linearity
plot(model, 2): check normality
plot(model, 3): check homoschedasticity
plot(model, 4): check influential cases


```{r}
#Checking assumptions of anova
plot(anova_log)

plot(anova_df)

plot(m2)
```

log transformation
```{r}
#Removing zeros, so we can do the log transformation. Need to argue that we can do that 
df_clean <- filter(df, df$suicides_no != "0")

#Making the log transformation
df_log <- df_clean %>% mutate(log_suicide = log(df_clean$suicide_population))

```


Trying to 
```{r}
m2 <- lm(df_log$log_suicide ~ df_log$age, data = df_log)

summary(m2)
```

```{r}
ggplot(df, aes(x = df$age, y = df$suicide_population)) +
  theme_minimal() +
    labs(x = "Age", y = "Suicide") +
  geom_bar(aes(fill= age), stat='summary', fun.y = mean, width = 0.5) +
  stat_summary(fun.data = mean_se, geom = "errorbar", color = 'black', width = 0.1) +
  ggtitle("The mean of suicide promille across ages")


ggplot(df, aes(x = df$age, y = df$suicides_no)) +
  theme_minimal() +
    labs(x = "Age", y = "Suicide") +
  geom_bar(aes(fill= age), stat='summary', fun.y = mean, width = 0.5) +
  stat_summary(fun.data = mean_se, geom = "errorbar", color = 'black', width = 0.1) +
  ggtitle("The mean of suicide across ages")
```

##### By visualizing the mean 
We can also support our judgement of results of ANOVA visually by drawing the mean values (just a reminder of code for boxplot with the mean value point). This will also allow to communicate results of the study in a more intuitive way.

*Draw a boxplot with the point for the mean for the second dataset based on example provided* 

```{r}
#Box plot

ggplot(df, aes(x = df$age, y = df$suicides_no, colour = age)) +
  geom_boxplot(width = 0.5) +
  ggtitle("Suicides across ages")+
  stat_summary(fun.y = mean, geom = "point", shape = 23, colour = "Black")
```

